---
title: "Getting Started with matchme - a package to facilitate matching mentees and mentors"
author: "Sarah Marshall"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with matchme - a package to facilitate matching mentees and mentors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

options(width = 150)
library(tidyverse)
library(matchme)
library(lpSolve)
```

# Introduction 

## Overview
## Data Requirements


# Example 1 - Extract "best match"

### Load data
```{r}

mentee_data <- read_csv("csvdata/example2_mentee_data.csv")

mentor_data <- read_csv("csvdata/example2_mentor_data.csv")


 
# create sample "original" matches (based on 1=1, 2=2 etc) - intentionally not a good match
sample_matches <- tibble(
  mentee_name = mentee_data$mentee_name,
  mentor_name = mentor_data$mentor_name[1:50]
)


```


### Specify matching criteria for 3 versions of the heuristic.

Criteria need to match column names of datasets.


"filters" allow only certain groups to be included in a particular iteration

```{r}

# version 1
# this has 11 iterations, each list has 11 elements
matching_criteria <- list()
  matching_criteria[[1]] <- tibble(
    mentee_criteria = list(c("mentee_gender_pref", "mentee_type_mentor_required", "mentee_programme", "mentee_major1"),
                           c( "mentee_type_mentor_required","mentee_programme", "mentee_major1"),
                           c("mentee_gender_pref", "mentee_type_mentor_required", "mentee_programme"),
                           c( "mentee_type_mentor_required","mentee_programme"),
                           c( "mentee_gender_pref", "mentee_type_mentor_required"),
                           c("mentee_type_mentor_required"),
                           c("mentee_gender_pref"),
                           c( "mentee_type_mentor_required","mentee_programme"),
                           c("mentee_type_mentor_required"),
                           c("mentee_programme"),
                           c()),
    mentor_criteria = list(c(  "mentor_gender",  "mentor_type", "mentor_programmes", "mentor_majors"),
                           c( "mentor_type", "mentor_programmes", "mentor_majors"),
                           c(  "mentor_gender",  "mentor_type", "mentor_programmes"),
                           c( "mentor_type", "mentor_programmes"),
                           c(  "mentor_gender","mentor_type"),
                           c("mentor_type"),
                           c("mentor_gender"),
                           c("mentor_type", "mentor_programmes"),
                           c( "mentor_type"),
                           c("mentor_programmes"),
                           c()),
    mentee_filter = list(NULL,  
                         "mentee_gender_pref == 'no_pref' ",
                         NULL,  
                         "mentee_gender_pref == 'no_pref' ",
                         NULL,
                         "mentee_gender_pref == 'no_pref' ",
                         NULL, NULL, NULL, NULL, 
                         NULL),
    mentor_filter = list(NULL, NULL, NULL, NULL, NULL, NULL, NULL, NULL, NULL, NULL, NULL)
  )
  
  
  
  
  matching_criteria[[2]] <- tibble(
    mentee_criteria = list(c( "mentee_gender_pref", "mentee_type_mentor_required", "mentee_programme"),
                           c( "mentee_type_mentor_required", "mentee_programme"),
                           c( "mentee_gender_pref", "mentee_type_mentor_required"),
                           c()),
    mentor_criteria = list(c( "mentor_gender", "mentor_type", "mentor_programmes"),
                           c( "mentor_type", "mentor_programmes"),
                           c(  "mentor_gender", "mentor_type"),
                          c()),
    mentee_filter = list(NULL,  
                         "mentee_gender_pref == 'no_pref' ",
                         NULL,
                        NULL),
    mentor_filter = list(NULL,NULL, NULL, NULL)
  )
  
  matching_criteria[[3]] <- tibble(
    mentee_criteria = list(c("mentee_gender_pref", "mentee_type_mentor_required", "mentee_programme"),
                           c("mentee_gender_pref"),
                           c()),
    mentor_criteria = list(c("mentor_gender", "mentor_type", "mentor_programmes"),
                           c("mentor_gender"),
                           c()),
    mentee_filter = list(NULL,  NULL,  NULL),
    mentor_filter = list(NULL, NULL, NULL)
  )
  
  

```

### Specify weights for LP
```{r}
# for LP only - each row is a different version.
# used with scoring_criteria to specify what columns should be used
# names in weights correspond to column names in mentee data
weights <- tribble(~gender_pref, ~type_mentor_required, ~programme,  ~major1, ~name,
                      0.5, 0.25, 0.15,0.1, "gender priority",
                     0.25, 0.25, 0.25,  0.25,  "equal",
                    1, 0, 0,0 ,  "gender only"
                     )





```

### Specify scoring criteria

* Criteria need to match column names of datasets.
* `_filters` allow pair to get a score if criteria is met, i.e. it overrules the mentee/mentor criteria.
* `match_type` specifies type of match. e.g.,
   - exact: value in column mentee_type_mentor_required must equal exactly the value in column mentor_type for a match to occur
   - subset: value in column mentee_programme must equal one of the values in column mentor_programmes_all for a match to occur
- exact with filter: value in column mentee_gender_pref must equal exactly the value in column mentor_gender, 
or mentee_gender_pref must equal  'no_pref' for a match to occur


```{r}

# used to compare all methods
scoring_criteria <- tibble(
  mentee_criteria = c("mentee_gender_pref", "mentee_type_mentor_required", "mentee_programme",  "mentee_major1"),
  mentor_criteria = c("mentor_gender",  "mentor_type", "mentor_programmes_all", "mentor_majors"),
  mentee_filter =  c("mentee_gender_pref == 'no_pref'", NA, NA, NA),
  mentor_filter = c(NA, NA,NA, NA),
  match_type = c("exact", "exact","subset", "subset"),
  weight = c(0.5,0.25, 0.15, 0.1)
) %>% mutate(
  match_name =  str_replace(mentee_criteria, "mentee_", ""))
 



```

### Get matches
```{r}

results <- get_and_evaluate_matches( mentee_data=mentee_data, mentor_data=mentor_data,
                        scoring_criteria = scoring_criteria,
                        weights = weights,
                        matching_criteria = matching_criteria,
                        valIfNA = 0,
                        compare_original = TRUE, 
                        original_matches_clean = sample_matches,
                        original_name = "alphabetical",
                        verbose = TRUE,
                        saveAllData = TRUE
                        )

```


Inspect performance of algorithms
```{r}
results$all_results
```

```{r, eval = TRUE}
best_matches = get_best_matches(results = results 
, criteria = "mean_score")

best_matches$matches
best_matches$unmatched_mentees
best_matches$unmatched_mentors

```

Notice that the unmatched mentors didn't complete most of the columns.

# Example 2 - Extract matches from all algorithms (small example + more details)

This example is a small example with 6 mentees and 7 mentors.

Matches are made with the following principles in mind:

1. Mentees are assigned a mentor of their preferred gender
1. Mentees are assigned a mentor of the appropriate type (1st year = academic, 2nd year +  = industry)
1. Mentees are assigned a mentor who aligns with the programme and subject of study.


### Load data
```{r}

mentee_data <- read_csv("csvdata/example1_mentee_data.csv")

mentor_data <- read_csv("csvdata/example1_mentor_data.csv")

# create sample "original" matches (based on 1=1, 2=2 etc) - intentionally not a good match
sample_matches <- tibble(
  mentee_name = mentee_data$mentee_name,
  mentor_name = mentor_data$mentor_name[1:6]
)
```


### Specify matching criteria for 2 versions of the heuristic.

Criteria need to match column names of datasets.


"filters" allow only certain groups to be included in a particular iteration

```{r}

# version 1
# this has 6 iterations, each list has 6 elements
matching_criteria <- list()
matching_criteria[[1]] <- tibble(
  mentee_criteria = list(c( "mentee_gender_pref", "mentee_type_mentor_required", "mentee_programme"),
                         c( "mentee_type_mentor_required", "mentee_programme"),
                         c( "mentee_gender_pref", "mentee_type_mentor_required"),
                         c("mentee_type_mentor_required"),
                         c("mentee_gender_pref"),
                         c()),
  mentor_criteria = list(c( "mentor_gender", "mentor_type", "mentor_programmes"),
                         c( "mentor_type", "mentor_programmes"),
                         c(  "mentor_gender", "mentor_type"),
                         c("mentor_type"),
                         c("mentor_gender"),
                         c()),
  mentee_filter = list(NULL,  
                       "mentee_gender_pref == 'no_pref' ",
                       NULL,
                        "mentee_gender_pref == 'no_pref' ",
                        NULL, 
                       NULL),
  mentor_filter = list(NULL, NULL, NULL, NULL, NULL, NULL)
)

# version 2
# this has 2 iterations, each list has 2 elements
matching_criteria[[2]] <- tibble(
  mentee_criteria = list(c("mentee_gender_pref", "mentee_type_mentor_required", "mentee_programme"),
                         c()),
  mentor_criteria = list(c("mentor_gender", "mentor_type", "mentor_programmes"),
                         c()),
  mentee_filter = list(NULL,    NULL),
  mentor_filter = list(NULL, NULL)
)




```


### Specify weights for LP
```{r}
# for LP only - each row is a different version.
# used with scoring_criteria to specify what columns should be used
# names in weights correspond to column names in mentee data

weights <- tribble(~gender_pref, ~type_mentor_required, ~programme,  ~major1, ~name,
                      0.5, 0.25, 0.15,0.1, "gender priority",
                     0.25, 0.25, 0.25,  0.25,  "equal",
                     )





```


### Specify scoring criteria

* Criteria need to match column names of datasets.
* `_filters` allow pair to get a score if criteria is met, i.e. it overrules the mentee/mentor criteria.
* `match_type` specifies type of match. e.g.,
   - exact: value in column mentee_type_mentor_required must equal exactly the value in column mentor_type for a match to occur
   - subset: value in column mentee_programme must equal one of the values in column mentor_programmes_all for a match to occur
- exact with filter: value in column mentee_gender_pref must equal exactly the value in column mentor_gender, 
or mentee_gender_pref must equal  'no_pref' for a match to occur



```{r}

# used to compare all methods
# criteria need to match 
scoring_criteria <- tibble(
  mentee_criteria = c("mentee_gender_pref", "mentee_type_mentor_required", "mentee_programme",  "mentee_major1"),
  mentor_criteria = c("mentor_gender",  "mentor_type", "mentor_programmes_all", "mentor_majors"),
  mentee_filter =  c("mentee_gender_pref == 'no_pref'", NA, NA, NA),
  mentor_filter = c(NA, NA,NA, NA),
  match_type = c("exact", "exact","subset", "subset"),
  weight = c(0.5,0.25, 0.15, 0.1)
) %>% mutate(
  match_name =  str_replace(mentee_criteria, "mentee_", ""))
 



```

### Get matches
```{r}

results <- get_and_evaluate_matches( mentee_data=mentee_data, mentor_data=mentor_data,
                        scoring_criteria = scoring_criteria,
                        weights = weights,
                        matching_criteria = matching_criteria,
                        valIfNA = 0,
                        compare_original = TRUE, 
                        original_matches_clean = sample_matches,
                        original_name = "alphabetical",
                        verbose = TRUE,
                        saveAllData = TRUE
                        )

```


Inspect performance of algorithms
```{r}

results$all_results
```


Inspect matches from 2 heuristics

```{r}
results$matches$heur_match[[1]]
results$matches$heur_match[[2]]

```


Inspect matches from 2 LP models

```{r}
results$matches$lp_match[[1]]$matches
results$matches$lp_match[[2]]$matches

```

Inspect original match data (in this case, alphabetical)

```{r}
results$matches$original_match

```

Additional information available about rule-based heuristic 

(example for 1st version of heuristic)

```{r}
results$matches$heur_match_all_data[[1]]$unmatched_mentors
results$matches$heur_match_all_data[[1]]$unmatched_mentees

```
```{r, results = "hide"}
results$matches$heur_match_all_data[[1]]$matches_by_iteration

```


# Example 3 - Non-student example

## Generate data

* mentee fields must start with "mentee_"
* mentee data must contain field "mentee_name"
* mentor fields must start with "mentor_"
* mentor data must contain field "mentor_name"


```{r}
set.seed(1)
n_mees = 100
n_mors = 90
fields = letters[1:5]
mbti = c("ISFJ", "ISTJ", "ESTJ", "ESFJ" , "ISFP", "ISTP", "ESTP", "ESFP", "INFJ", "INTJ", "ENTJ", "ENFJ", "INFP", "INTP", "ENTP", "ENFP")

# properties must start with "mentee_"
(mentee_data <- tibble(
  mentee_name = paste0("A", 1:n_mees), # must have mentee_name field
  mentee_field = sample(size = n_mees,  fields, replace = TRUE),
  mentee_mbti = sample(size = n_mees, c("no pref",mbti), replace = TRUE)
))

mentor_data <- tibble(
  mentor_name = paste0("B", 1:n_mors), # must have mentor_name field
  mentor_field = sample(size = n_mors, fields, replace = TRUE),
  mentor_mbtiiii = sample(size = n_mors, mbti, replace = TRUE)
) %>% rowwise() %>% 
  mutate( 
    mentor_fields = paste(unique(c(mentor_field, sample(size = 2, fields, replace = TRUE))),   collapse = ";")
)
 

```


### Specify matching criteria for 2 versions of the heuristic.

Criteria need to match column names of datasets.


"filters" allow only certain groups to be included in a particular iteration

```{r}

# version 1
# this has 6 iterations, each list has 6 elements
matching_criteria <- list()
matching_criteria[[1]] <- tibble(
  mentee_criteria = list(c( "mentee_field", "mentee_mbti"),
                         c( "mentee_field"),
                         c( "mentee_mbti"),
                         c()),
  mentor_criteria = list(c( "mentor_field", "mentor_mbtiiii"),
                         c( "mentor_field"),
                         c(  "mentor_mbtiiii"),
                         c()),
  mentee_filter = list(NULL,  
                       "mentee_mbti == 'no_pref'",
                       NULL, 
                       NULL),
  mentor_filter = list(NULL, NULL, NULL, NULL)
)

# version 2
# this has 2 iterations, each list has 2 elements
matching_criteria[[2]] <- tibble(
  mentee_criteria = list(c("mentee_field"),
                         c()),
  mentor_criteria = list(c("mentor_field"),
                         c()),
  mentee_filter = list(NULL,    NULL),
  mentor_filter = list(NULL, NULL)
)




```

### Specify weights for LP
```{r}
# for LP only - each row is a different version.
# used with scoring_criteria to specify what columns should be used
# names in weights correspond to column names in mentee data
weights <- tribble(~field, ~mbti, ~name,
                      0.75, 0.25, "field priority",
                     0.5, 0.5 , "equal",
                     )





```

### Specify scoring criteria

* Criteria need to match column names of datasets.
* `_filters` allow pair to get a score if criteria is met, i.e. it overrules the mentee/mentor criteria.
 

```{r}

# used to compare all methods
scoring_criteria <- tibble(
  mentee_criteria = c("mentee_field", "mentee_mbti"),
  mentor_criteria = c("mentor_fields",  "mentor_mbtiiii"),
  mentee_filter =  c(NA, "mentee_mbti == 'no pref'"),
  mentor_filter = c(NA, NA),
  match_type = c("subset", "exact"),
  weight = c(0.75,0.25)
) %>% mutate(
  match_name =  str_replace(mentee_criteria, "mentee_", ""))
 



```

### Get matches
```{r}

results <- get_and_evaluate_matches( mentee_data=mentee_data, mentor_data=mentor_data,
                        scoring_criteria = scoring_criteria,
                        weights = weights,
                        matching_criteria = matching_criteria,
                        valIfNA = 0,
                        compare_original = TRUE, 
                        #original_matches_clean = sample_matches,
                        #original_name = "alphabetical",
                        verbose = TRUE,
                        saveAllData = TRUE
                        )

```


Inspect performance of algorithms
```{r}
results$all_results

#results$matches$heur_match[[1]]%>% arrange(mentee_name)

```

```{r, eval = TRUE}
best_matches = get_best_matches(results = results 
, criteria = "mean_score")

best_matches$matches %>% arrange(desc(mentee_mbti))
best_matches$unmatched_mentees
best_matches$unmatched_mentors

```

